\singlespacing{\textbf{Efficient pollution abatement in electricity markets with intermittent renewable energy}}

### 2019-05-10 documents/draft.tex
Update draft
- abstract clean up
- updates to method
- rewrite beginning of discussion
- label subsubsections
- add VES approx proof
- add econometric methodology explanation

\documentclass[11pt,a4paper]{extarticle}
\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{booktabs} % for toprule, midrule and bottomrule
\usepackage{adjustbox}
\usepackage{amsmath}
\usepackage{bbold}
\usepackage{etoolbox}
\usepackage{setspace} % for \onehalfspacing and \singlespacing macros
\usepackage[hidelinks]{hyperref}
\usepackage{array}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{caption}
\usepackage{pdflscape}
\usepackage{caption}
\usepackage{tabularx}
\usepackage{authblk}
\usepackage{float}
\usepackage{siunitx}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage[authoryear]{natbib}
\usepackage{scrextend}
\usepackage{nicefrac}
\usepackage{enumitem}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{cleveref}
\usepackage{varwidth}
%\usepackage{showframe}
%\usepackage{lipsum}

@@

% Syntax: \begin{addmargin}[<left indentation>]{<indentation>}
\begin{addmargin}[0.5in]{0.5in}
	\textit{ Use of renewable energy has increased greatly ... 
		We evaluate renewable energy technologies in a partial equilibrium setting that integrates their production profiles with variation in the market price of electricity. 
		In order to discount the ..., we use a representative agent that prefers to smooth their electricity consumption over time. 
		We find a non-linear elasticity of substitution between intermittent and non-intermittent generation technologies. Hence, a CES production function fails to accurately model substitutability between renewables and fossil fuels. Alternatively, we find that a VES production function can reasonably approximate the effect of intermittency on the elasticity of substitution between between intermittent and non-intermittent energy. 
		Furthermore, our results suggest that the welfare burden of carbon taxes on consumers will vary geographically based on the availability of consistent, renewable technology. This distributional side effect can be counteracted by subsidizing research into battery technologies; improvements in batteries can also complement present policy to promote renewables by greatly increasing the substitutability of intermittent  and non-intermittent energy technologies. 
	 }
\end{addmargin}

@@

Renewable energy technologies have seen considerable adoption over the last few decades (EIA 2019). Unlike the alternatives, wind and solar power are particularly unique in that the amount of energy they supply is intermittent. Consequently, designing economically efficient policy to promote their adoption is not straightforward given that they cannot easily substitute for fossil fuel technologies such as coal power. Moreover, traditional economic measures such as the levelized cost of electricity (LCOE) fail to capture the true economic value of intermittent technologies, because they neglect to account for variation in output and prices over time \citep{Joskow2011}.

Some of the literature has overcome this critique by constructing numerical models that find the cheapest renewable technology set while accounting for intermittent supply. For instance, \citet{MN2006} model uncertain renewable output with intertemporal generation constraints, while \citet{NCK2007} model temporal and spatial characteristics of wind output to optimize its deployment in the UK. On the other hand, other literature focuses on the effect of intermittent technologies on the market itself; \citet{AC2012} study the interaction between intermittent renewables and traditional reliable sources of energy in decentralized markets, and \citet{Chao2011} models alternative pricing mechanisms for intermittent renewable energy sources. Additionally, \citet{Boren2012} reviews the effects of present public policies used to promote renewables and the challenges posed by intermittency.

Our model comes closest to that of \citet{HH} who build a peak-load pricing model where the availability of renewable capacity varies stochastically. This stochastic variability models intermittency and negatively affects the adoption of renewables. Furthermore, they simplify their analysis by setting aside complications such as  outage costs and rationing rules. In total, their model finds an S-shaped adoption curve for renewables as they get cheaper. In addition, they find that a Pigouvian tax can properly internalize the costs of fossil fuels in a setting with perfect competition. Likewise, we model the  equilibrium of a market with dynamic prices and access to both renewable and fossil energy. However, our results differ in key ways, because we take a different approach to modeling intermittency.

To highlight this difference, we must first contrast the terms  \textit{intermittency} and \textit{reliability}. By intermittency, we predictable changes in output related to physical constraints. For instance, while wind energy output varies over time, if  we know the angle and speed of wind, we may precisely derive the quantity of energy that a wind farm generates.\footnote{ \citet{Foley2012} provide a review of the literature on the forecasting of wind power. Over time, forecasts have gotten much more accurate allowing electricity grids to manage wind power intermittency ahead of time.}  At the same time, these forecasts may not be perfect and, as a result, we see unexpected variation in the output of a technology. This unpredictability may be better captured by the notion of reliability; specifically,  we refer to definition provided by the US Department of Energy's \citet{ORNL} -- ``Power reliability can be defined as the degree to which the performance of the elements in a bulk system results in electricity being delivered to customers within accepted standards and in the amount desired." This is more so related to the stochastic or unpredictable variation in the output of a technology. Unlike intermittency, this cannot be planned around with 100\% certainty. A practical example may be a wind turbine's systems failing. While we may know the chances of this occurring, it's not possible to know when it will occur ahead of time; consequently, this may result in a temporary reduction in the quantity and quality of the electricity delivered --  a loss of reliability. In short, when \citeauthor{HH} define renewable output as equal to a baseline capacity multiplied by a uniform random variable, their model is closer to one of reliability rather than intermittency.

On the other hand,  we model intermittency as defined above by allowing the output of renewable energy to differ between periods according to known function. For parsimony, we assume that the output of all technologies in each period does not vary stochastically; that is, we focus on intermittency rather than reliability. Next, we model the electricity sector using a representative firm that chooses and builds capacity from a set of electricity-generating technologies to maximize profit; some of these technologies are intermittent while others have constant output. Then, we consider a representative consumer who purchases varying quantities of electricity in each period in order to maximize utility. Generally, people prefer to spread their electricity consumption out over time. We model this preference by using a CES function of electricity consumption differentiated by period.\footnote{ The use of a CES function in this way has been explored earlier by \citet{Schwarz}, \citet{Schwarz}, \citet{Herriges}, and \citet{KS1994}, \citet{Aubin}, and \citet{Moha2016}. Their papers empirically estimate the parameters for this function; we provide a more detailed discussion of the empirical literature in our \hyperref[sec:methodology]{Methodology}.} These two sides of the market reach an equilibrium through adjustments in prices of electricity in each period. All in all, we make many of the same assumptions that \citet{HH} do; we assume dynamic pricing,\footnote{ \citeauthor{HH} provide a motivation for models incorporating the dynamic pricing of electricity. They argue such approaches to pricing will become the norm with further technological advances and coming regulatory changes.} no load rationing, and positive prices.

Then, we parametrize our model empirically. To start, we fit the parameters of the consumer's CES utility function using electricity consumption and price data for each US state. Specifically, we estimate the intertemporal elasticity of substitution for electricity consumption; this parameter plays a particularly important role in our model, since it captures the effects of intermittency on demand. Then, to model the supply side, we narrow our model to a two-period, two-technology setting with renewables and fossil energy. We proxy for renewable energy using solar and proxy for fossil energy using coal; then, we parametrize each accordingly. Finally, we implement our model numerically and solve for the equilibrium. We study the effects of technical change on the energy market and derive policy implications.

Overall, this treatment of intermittency produces some key results that contrast with those of Helm and Mier. First, while we also find that externalities can be handled through a Pigouvian tax, we further find that a combining both taxes on pollution and research subsidies for renewables is more optimal than either policy instrument alone. Second, we find that, as renewables get cheaper, their adoption is not necessarily S-shaped but can be concave or convex depending the elasticity of substitution in the consumer's utility function; with our estimated parameters, we find this relationship to be almost linear. Third, a significant amount of literature has assumed a CES structure between renewable and fossil energy (see \citet{Pap}); this assumption has been motivated, in part, by the need to capture imperfect substitutability between these two generation technologies as a result of intermittency. However, our direct model of intermittency disconfirms this assumption; the elasticity of substitution between these two energy technologies is far from constant as a \textit{consequence} of intermittency. Moreover, it varies with the prices and quantities of energy generation technologies employed. Hence, we believe that optimal environmental policy, such as pollution taxes and research subsidies, should take into account the technologies powering local energy markets in order to effectively handle externalities.

@@

Consumers purchase variable amounts of electricity $Z_t$ over each period $t$. Additionally, assuming that the price of electricity in each period is held constant, they prefer to ration their electricity usage across each period in fixed proportions. And, lastly, we presume that consumers would be willing to shift their consumption from one period to another in response to a shift in prices. Overall, these assumptions can be captured using a a standard CES utility function
\begin{equation}
U = \left( \sum_t \alpha_t Z_t^\phi  \right)^{1/\phi}
\end{equation}
where $\sigma = 1/(1-\phi)$ is the intertemporal elasticity of substitution for electricity consumption. We define $\sum_t \alpha_t = 1$, so that $\alpha_t$ is the fraction of electricity consumption in period $t$ when all prices are equal; naturally, $\alpha_t > 0$ for all $t$. We assume that any discounting over time is absorbed into $\alpha$ for parsimony.The budget constraint is given by
\begin{equation}
I = \sum_t p_t Z_t
\end{equation}
where $p_t$ is the price of electricity in period $t$ and $I$ is the income. Furthermore, since CES preferences are homothetic, we may aggregate the consumers into a single representative consumer. Consequently, the first order conditions four our representative consumer when maximizing utility given this budget constraint imply:
\begin{align}
Z_t &= \left(\frac{\alpha_t}{p_t} \right)^\sigma \frac{I}{P} \\
P &= \sum_t \alpha_t^\sigma p_t^{1-\sigma}
\end{align}
where $P$ is the price index. Note that this model naturally does not allow for blackouts in equilibrium, since the price of electricity in any period gets arbitrarily large as the quantity of energy consumed in that period approaches 0. Furthermore, note that prices must be positive; although this is sometimes violated in reality, we do not believe that this assumption significantly affects our analysis.

@@

Secondly, we have firms maximizing profit by picking an optimal set of energy inputs. In reality, electricity markets are fairly competitive, so we can model the set of firms by using a single representative firm that sets marginal revenue equal to marginal cost. We define the quantity of deployed energy technology $i$ as $X_i$, and we define its output per unit in period $t$ as $\xi_{1,t}$. So, for example, if $i$ is solar power, $X_i$ would be the number of solar panels and $\xi_{i,t}$ may be kW generated per solar panel in period $t$. Consequently, the energy generated in period $t$, $Z_t$, is given by $\sum_i \xi_{i,t} X_i$. We define these variables in matrices to simplify notation:
\begin{align*}
X \equiv \begin{pmatrix}
X_1\\
\vdots\\
X_n
\end{pmatrix} ,\;
Z \equiv \begin{pmatrix}
Z_1\\
\vdots\\
Z_m
\end{pmatrix} ,\;
p \equiv \begin{pmatrix}
p_1\\
\vdots\\
p_m
\end{pmatrix} ,\;
\xi \equiv \begin{pmatrix}
\xi_{1,1} & \dots & \xi_{1,m}\\
\vdots & \ddots & \\
\xi_{n,1} &  & \xi_{n,m}
\end{pmatrix} 
\end{align*}
where we have $n$ technologies, $m$ periods, and $Z \equiv \xi^T X$. A key element of this model is that $X$ does not vary by time. This is important, because the time scale of our model must be fairly granular to study the effects of intermittency. So, for instance, we may have $t \in \{ 1, \dots, 24\}$ representing each hour of the day. In such short time scales, we can safely assume that producers do not modify the quantity of the technology they employ. Moreover, this keeps the model parsimonious by preventing more complicated dynamics from entering the producers optimization problem.

Instead, our representative firm sets $X$ once to maximize profit in all periods while facing the cost function $C(X_1, \dots, X_n)$.  \citet{HH} note that past literature has argued for concave cost functions in the energy market due to effects such as economics of scale and learning by doing; on the other hand, standard cost functions are generally convex. So, like Helm and Meir, we take an intermediate approach by using a linear cost function. Specifically, we have $C(X) \equiv \sum_i c_i X_i \equiv c^T X$ where $c_i$ is the cost per unit of $X_i$. Total profit is given by 
\begin{equation}
\Pi = p^T Z - c^T X
\end{equation}
To simplify the algebra, we set the number of technologies equal to the number of periods $(n=m)$. Additionally, we further require that the output per unit of each technology is unique and non-negative in each period; in other words, the output per unit of one technology is not a linear combination of those of the other technologies in our set. This then implies that $\xi$ is of full rank and therefore invertible.  Now, maximizing profit, we find the first order condition:
\begin{equation}
\frac{\partial \Pi}{\partial X} = 0 \implies p = \xi^{-1} c
\end{equation}
Combining this FOC with the demand equation (Equation (3)) allows us to find the equilibrium. Generally, the equilibrium results for any number of technologies ($n$) are analytic, but they are difficult to interpret due to the number of parameters involved.

@@

For more tractable results, we consider a simpler scenario where $n = m = 2$ and $\sigma = 1$ (Cobb-Douglas); this particular case is described in greater  detail with proofs in \hyperref[sec:cobbdoug]{Appendix A.A}. Moreover, these parameters are of particular interest, because they simplify the model enough to allow us to derive comparative statics and determine which sets of parameters lead to edge cases.

To start, we discuss the conditions on the exogenous parameters required to avoid edge-cases. Alternatively, these are the conditions required for a technology to be economical. But, first, we define two recurring terms in our analysis: cost efficiency and output efficiency. For some arbitrary period $t$ and technology $i$, we use cost efficiency to refer to $\xi_{i,t}/c_i$ and output efficiency to refer to $\xi_{i,t}$. So, for example, A technology is more cost efficient than another in a particular period when its output in that period per ``dollar" spent is larger than that of the other technology in the same period. 
\begin{proposition}
	Assume that, for all  $i$ (technologies) and  $t$ (periods), we have $\xi_{i,t} > 0 \, , \alpha_t > 0 \, ,$ and $ c_i > 0$. Furthermore, suppose that $\xi$ is invertible. For technology $i$ to be economical, it must meet three conditions with respect to some arbitrary period $t$. Firstly, it must the more cost effective than any of the other technologies in period $t$. Secondly, it must have have a comparative advantage in output efficiency in the same period $t$. And, thirdly, there must be sufficient amount of demand for energy in this period $t$. 
\end{proposition}
The condition on cost efficacy is fairly intuitive; one case of the contrapositive is that if a technology is not the most cost effective in any period, it will not be used. Alternatively, if a technology is the most cost effective in every single period, it will be the only technology used. The second condition regarding output efficiency actually stems from the invertibility of $\xi$. If we did not have $\xi$ invertible, we either have at least one technology that does not produce in any period or we have at least one technology being a linear combination of the other technologies in terms of output. In the latter case, some technologies are not economical because their output can be replicated the other technologies in a more cost effective way. As an example, suppose $\xi$ consisted of 3 technologies but was of rank $2$. In this case, we may represent any of these technologies as a linear combination of the other two; call this combination the synthetic version of technology $i$. It is not possible for every synthetic technology $i$ to be more expensive than its original; for at least one $i$, we must have a synthetic technology being cheaper or equal in price to the actual technology. In the first case, synthetic technology $i$ is cheaper so we do not use the actual technology $i$. In the later case, we may still eliminate technology $i$ or not use the other two technologies depending on the demand. Overall, in any case, $\xi$ not invertible means that at least one technology is not used. Finally, the demand condition is straightforward; even if a technology is optimal in a certain period, if consumers do not sufficiently demand electricity during that period, then there is little reason to use that technology.

We may also derive the comparative statics for this simplified scenario.  
\begin{proposition}
	Suppose we are not in an edge case, so that the conditions of Proposition 1 hold for each technology. The equilibrium quantity of a technology is increasing with its output efficiency and decreasing with its cost; at the same time, it is decreasing with the output efficiency of the other technologies and increasing with their cost. Also, suppose that some technology $i$ is the most cost effective in period $t$. Then, its equilibrium quantity is increasing with respect to the demand share parameter $\alpha_t$ and decreasing with respect to the demand share parameters of the other periods. Furthermore, again assuming technology $i$ is the most cost effective in period $t$, the comparative statics of $Z_t$ and $X_i$ are equivalent. 
\end{proposition}	
The comparative statics with respect to $X$ and its output efficiency and cost are not surprising. On the other hand, the statics for $Z$ may not be immediately obvious. Instead, they follow from the fact that we have $Z \equiv \xi_T X$. That is, suppose we have an arbitary technology $i$ that is the most cost effective source of electricity on period $t$. If consumers demanded that 100\% of their energy arrive in period $t$, then relying solely on technology $i$ for energy would be the most economical solution. Consequently, it seems intuitive that the comparative statics of $X_i$ follow through to $Z_t$. This intuition just happens to apply even when other technologies are employed and there is demand in multiple periods. Similarly, the comparative statics for the share parameters of the utility function, $\alpha$, travel in the opposite direction. A rise in $\alpha_t$ would directly raise the optimal quantity of $Z_t$; hence, whichever technology is most cost effective at producing in period $t$ would be used more. We provide a more detailed and quantitative discussion of the comparative statics and edge cases in the appendix.

@@

\section{Results}

@@

\subsection{Implications for Pollution Abatement}

Overall, our results have shown that the traditional assumption of a CES relationship between renewable and fossil energy is inappropriate. However, since $e$ takes on a fairly simple shape, we can still adjust the findings of past models to account for variation in the elasticity of substitution. For example, consider again the results of \citet{Ace2012}.

>>>>

\documentclass[11pt,a4paper]{extarticle}
\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{booktabs} % for toprule, midrule and bottomrule
\usepackage{adjustbox}
\usepackage{amsmath}
\usepackage{bbold}
\usepackage{etoolbox}
\usepackage{setspace} % for \onehalfspacing and \singlespacing macros
\usepackage[hidelinks]{hyperref}
\usepackage{array}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{caption}
\usepackage{pdflscape}
\usepackage{caption}
\usepackage{tabularx}
\usepackage{authblk}
\usepackage{float}
\usepackage{siunitx}
\usepackage{titlesec}
\usepackage{pgfplots}
\usepackage[authoryear]{natbib}
\usepackage{scrextend}
\usepackage{nicefrac}
\usepackage{enumitem}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{cleveref}
\usepackage{varwidth}
\usepackage{gensymb}
%\usepackage{showframe}
%\usepackage{lipsum}

@@

% Syntax: \begin{addmargin}[<left indentation>]{<indentation>}
\begin{addmargin}[0.5in]{0.5in}
	\textit{In this paper, we present a model of the electricity sector where generation technologies are intermittent. The economic value of a electricity generation technology is given by integrating its production profile with the market price of electricity. We parametrize this model empirically and generate numerical results. The introduction of intermittency results in a non-constant elasticity of substitution between renewable and fossil energy. Consequently, this suggests that the efficacy and welfare effects of carbon taxes and renewable subsidies vary geographically. Subsidizing research into battery technology can mitigate this distributional side effect while complementing traditional policies used to promote renewable energy. 
	 }
\end{addmargin}

@@

Renewable energy technologies have seen considerable adoption over the last few decades \citep{EIArenew}. Unlike the alternatives, wind and solar power are particularly unique in that the amount of energy they supply is intermittent. Consequently, designing economically efficient policies to promote their adoption is not straightforward given that they cannot easily substitute for fossil fuel technologies such as coal power. Moreover, traditional economic measures such as the levelized cost of electricity (LCOE) fail to capture the true economic value of intermittent technologies, because they neglect to account for variation in output and prices over time \citep{Joskow2011}.

Some of the literature has approached this problem by constructing numerical models that find the cheapest renewable technology set while accounting for intermittent supply. For instance, \citet{MN2006} model uncertain renewable output with intertemporal generation constraints, while \citet{NCK2007} model the temporal and spatial characteristics of wind output to optimize its deployment in the UK. Other papers study how intermittent technologies affect the market itself; \citet{AC2012} study the interaction between intermittent renewables and traditional reliable sources of energy in decentralized markets, and \citet{Chao2011} models alternative pricing mechanisms for intermittent renewable energy sources. Additionally, \citet{Boren2012} reviews the effects of present public policies used to promote renewables and the challenges posed by intermittency.

Our model comes closest to that of \citet{HH} who build a peak-load pricing model where the availability of renewable capacity varies stochastically. This stochastic variability models intermittency and negatively affects the adoption of renewables. Furthermore, they simplify their analysis by setting aside complications such as  outage costs and rationing rules. Their model finds an S-shaped adoption curve for renewables as they get cheaper. In addition, they find that a Pigouvian tax can properly internalize the costs of fossil fuels in a setting with perfect competition. Like \citeauthor{HH}, we model the  equilibrium of a market with dynamic prices and access to both renewable and fossil energy. However, our results differ in key ways, because we take a different approach to modeling intermittency.

To highlight this difference, we must first contrast the terms  \textit{intermittency} and \textit{reliability}. By intermittency, we mean predictable changes in output related to physical constraints. For instance, while wind energy output varies over time, if  we know the wind speed and angle, we may precisely derive the quantity of energy that a wind farm generates.\footnote{ \citet{Foley2012} provide a review of the literature on the forecasting of wind power. Over time, forecasts have gotten much more accurate allowing electricity grids to manage wind power intermittency ahead of time.}  At the same time, wind forecasts are not be perfect, so we may see unexpected variation in the electricity output of a wind farm. This unpredictability is better captured by the notion of reliability; specifically,  we refer to definition provided by the US Department of Energy's \citet{ORNL} -- ``Power reliability can be defined as the degree to which the performance of the elements in a bulk system results in electricity being delivered to customers within accepted standards and in the amount desired." This is more so related to the stochastic or unpredictable variation in the output of a technology. Unlike intermittency, this cannot be planned around with 100\% certainty. A practical example may be a wind turbine's systems failing. While we may know the chances of this occurring, it's not possible to know when it will occur ahead of time; consequently, this may result in a temporary reduction in the quantity and quality of the electricity delivered --  a loss of reliability. In short, when \citeauthor{HH} model renewable output as equal to a base capacity multiplied by a uniform random variable, their treatment is closer to one of reliability rather than intermittency.

On the other hand,  we model intermittency as defined above by allowing the output of renewable energy to differ between periods according to known function. For parsimony, we assume that the output of each technology  does not vary stochastically; that is, we do not study reliability. Next, we model the electricity sector using a representative firm that chooses and builds capacity from a set of electricity-generating technologies to maximize profit; some of these technologies are intermittent while others have constant output. Then, we consider a representative consumer who purchases varying quantities of electricity in each period in order to maximize utility. People prefer to smooth their electricity consumption over time, so we model our consumer's preferences using a CES function of electricity consumption differentiated by period.\footnote{ The use of a CES function in this way has been explored earlier by \citet{Schwarz}, \citet{Schwarz}, \citet{Herriges}, and \citet{KS1994}, \citet{Aubin}, and \citet{Moha2016}. Their papers empirically estimate the parameters for this function; we provide a more detailed discussion of the empirical literature in our \hyperref[sec:methodology]{Methodology}.} These two sides of the market reach an equilibrium through adjustments in prices of electricity in each period. Overall, we make many of the same assumptions that \citet{HH} do; we assume dynamic pricing,\footnote{ \citeauthor{HH} motivate models incorporating the dynamic pricing of electricity. They argue that such approaches to pricing will become the norm with further technological advances and coming regulatory changes.} no load rationing, and positive prices.

Then, we parametrize our model empirically. We fit the parameters of the consumer's CES utility function using electricity consumption and price data for each US state. That is, we estimate the intertemporal elasticity of substitution for electricity consumption; this parameter plays a particularly important role in our model, since it captures the effects of intermittency on demand. Next, to model the supply side, we narrow our framework to a two-period, two-technology setting to focus on the substitutability between renewable and fossil energy. We proxy for renewable and fossil energy using solar and coal; we parametrize each accordingly. Finally, we implement our model numerically and provide suggestions for policy and future models.

Our results show that the elasticity of substitution between renewable and fossil energy is not constant. This is important because a significant amount of literature has assumed a CES structure between renewable and fossil energy (see \citet{Pap}); this assumption has been motivated, in part, by the need to capture imperfect substitutability between these two generation technologies as a result of intermittency. But, our model finds that the elasticity of substitution between renewable and fossil energy is far from constant as a \textit{consequence} of intermittency. Specifically, this elasticity appears to vary with the quantity of each energy technology. Hence, we argue against assuming a CES structure. Alternatively, we find that a VES production function can be used to approximate the relationship that our model finds between the elasticity of substitution and the quantities of each technology. That is, this relationship is roughly linear which is exactly what a VES function assumes. Moreover, a VES function is analytically tractable, so it can be implemented in other frameworks without making them overly complicated.

Additionally, our results have multiple important implications for environmental/energy policy. Firstly, we find that the welfare burden of a carbon tax  and the effect of subsidies on renewable adoption both vary geographically. This variation is a consequence of differences in the intermittency and availability of renewable energy by location. Moreover, this can create a trade-off between equitably and efficiently mitigating climate change. Hence, policymakers should account for geographic variation in energy markets to ensure they handle this equity-efficiency trade-off appropriately. Secondly, we find strong motivation for research subsidies aimed at improving battery technology. That is, better energy storage can greatly increase the elasticity of substitution between renewable and fossil energy.  Also, batteries can  lessen the distributional side effects of carbon taxes and renewable subsidies by  reducing intermittency. Hence, research into improving batteries can complement other policies by making them more equitable and by increasing their impact on the adoption of renewables. Finally, we revisit the results of \citet{Ace2012} and qualitatively discuss the implications of our model in their setting. We now discuss our model in greater detail.

@@

Consumers purchase a quantity of electricity $Z_t$ in each period $t$. Furthermore, they demand a greater quantity of electricity in certain periods; for instance, consumers need more electricity during the middle of the day more than at night. At the same time, consumers are willing to shift their consumption from one period to another in response to a shift in prices. Overall, these characteristics can be captured using a standard CES utility function.  Furthermore, since CES preferences are homothetic, we may aggregate the consumers into a single representative consumer. This consumer has the utility function
\begin{equation}
U = \left( \sum_t \alpha_t Z_t^\phi  \right)^{1/\phi}
\end{equation}
where $\sigma = 1/(1-\phi)$ is the intertemporal elasticity of substitution for electricity consumption. To ensure that indifferences surfaces are convex, we have $\phi \in (0,1)$ or, equivalently,  $\sigma > 0$; additionally, since electricity consumption increases utility, we must have $\alpha_t > 0$ for all $t$.  For simplicity, we also define $\sum_t \alpha_t = 1$, so that a 1\% increase in electricity consumption in all periods increases utility by 1\%; consequently, we then have $\alpha_t \in (0,1)$. The budget constraint is given by
\begin{equation}
I = \sum_t p_t Z_t
\end{equation}
where $p_t$ is the price of electricity in period $t$ and $I$ is the income. Our representative consumer maximizes utility against this budget constraint; the first order conditions of this problem imply:
\begin{align}
Z_t &= \left(\frac{\alpha_t}{p_t} \right)^\sigma \frac{I}{P} \\
P &= \sum_t \alpha_t^\sigma p_t^{1-\sigma}
\end{align}
where $P$ is the price index. Note that this model naturally does not allow for blackouts in equilibrium, since the price of electricity in any period gets arbitrarily large as the quantity of energy consumed in that period approaches 0. Furthermore, note that prices must be positive; although this is sometimes violated in reality, we do not believe that this assumption significantly affects our analysis.

@@

Secondly, we have firms maximizing profit by picking an optimal set of energy inputs. In reality, electricity markets are fairly competitive, so we can model the set of firms by using a single representative firm that sets marginal revenue equal to marginal cost.

We let $X_i$  represent the quantity of energy technology $i$ , and we define its output per unit in period $t$ as $\xi_{1,t}$. So, for example, if $i$ is solar power, $X_i$ would be the number of solar panels and $\xi_{i,t}$ may be kWh generated per solar panel in period $t$. Consequently, the energy generated in period $t$, $Z_t$, is given by $\sum_i \xi_{i,t} X_i$. To simplify notation, we have
\begin{align*}
X \equiv \begin{pmatrix}
X_1\\
\vdots\\
X_n
\end{pmatrix} ,\;
Z \equiv \begin{pmatrix}
Z_1\\
\vdots\\
Z_m
\end{pmatrix} ,\;
p \equiv \begin{pmatrix}
p_1\\
\vdots\\
p_m
\end{pmatrix} ,\;
\xi \equiv \begin{pmatrix}
\xi_{1,1} & \dots & \xi_{1,m}\\
\vdots & \ddots & \\
\xi_{n,1} &  & \xi_{n,m}
\end{pmatrix} 
\end{align*}
where we have $n$ technologies, $m$ periods, and $Z \equiv \xi^T X$. A key element of this model is that $X$ does not vary by time. This is important because the time scale of our model must be fairly granular to study the effects of intermittency. So, for instance, we may have $t \in \{ 1, \dots, 24\}$ representing each hour of the day. Over such short time scales, the intermittency of renewables like solar and wind can significantly affect their economic value. Additionally, this allows us to assume that producers cannot modify the quantity of the technologies they deploy. And, thirdly, for certain technologies, this time frame is short enough to assume that $\xi$ is exogenous; that is, technologies like coal power cannot significantly modify their output within a day, so $\xi$ can be treated as a given set of constants. These latter two assumptions, that $X$ cannot change over time and that $\xi$ is exogenous, are important for parsimony, because they prevent more complicated dynamics for entering our model. So, by assuming a sufficiently short time frame, we can continue our study of intermittency in a relatively simpler setting.

Next, our representative firm sets $X$ once to maximize total profit while facing the cost function $C(X_1, \dots, X_n)$.  \citet{HH} note that past literature has argued for concave cost functions in the energy market due to effects such as economies of scale and learning-by-doing; on the other hand, standard cost functions are generally convex. So, like Helm and Meir, we take an intermediate approach by using a linear cost function. Specifically, we have $C(X) \equiv \sum_i c_i X_i \equiv c^T X$ where $c_i$ is the cost per unit of $X_i$. Total profit is given by 
\begin{equation}
\Pi = p^T Z - c^T X
\end{equation}
To simplify the algebra, we set the number of technologies equal to the number of periods $(n=m)$. Additionally, we further require that the output per unit of each technology is unique and non-negative in each period; in other words, the output per unit of one technology is not a linear combination of those of the other technologies in our set. This then implies that $\xi$ is of full rank and therefore invertible.  Now, maximizing profit, we find the first order condition:
\begin{equation}
\frac{\partial \Pi}{\partial X} = 0 \implies p = \xi^{-1} c
\end{equation}
Combining this FOC with the demand equation (Equation (3)) allows us to find the equilibrium. Generally, the equilibrium results for any number of technologies ($n$) are analytic, but they are difficult to interpret due to the number of parameters involved.

@@

For more tractable results, we consider a simpler scenario where $n = m = 2$ and $\sigma = 1$ (Cobb-Douglas); this particular case is described in greater  detail  in \hyperref[sec:cobbdoug]{Appendix A.A}. These parameters simplify the model enough to allow us to derive comparative statics and determine which sets of parameters lead to edge cases. Furthermore, as we will discuss in our \hyperref[sec:results]{results}, assuming $\sigma=1$ is not far from its empirical estimate.

To start, we discuss the conditions required on the exogenous parameters  to avoid edge-cases. But  first, we define two recurring terms in our analysis: cost efficiency and output efficiency. For some arbitrary period $t$ and technology $i$, we use cost efficiency to refer to $\xi_{i,t}/c_i$ and output efficiency to refer to $\xi_{i,t}$. So, for example, with two technologies $i$ and $j$ and an arbitrary period $t$, technology $i$ is more cost efficient than technology $j$ in period $t$ when $\xi_{i,t}/c_i > \xi_{j,t}/c_j$. Additionally, we say a technology is economical if its equilibrium quantity is above 0. In order to avoid an edge-case, we require both technologies to be economical; we give the conditions to ensure this for each technology in Proposition 1. 
\begin{proposition}
	Assume that, for all technologies $i$ and  periods $t$, we have $\xi_{i,t} > 0 \, , \alpha_t > 0 \, ,$ and $ c_i > 0$. For technology $j$ to be economical, it must meet three conditions with respect to some arbitrary period $s$. Firstly, it must be more cost effective than any of the other technologies in period $s$. Secondly, it must have  a relative advantage in output efficiency in the same period $s$. And, thirdly, there must be sufficient amount of demand for energy in this period $s$. 
\end{proposition}
The first condition is on cost efficacy and is fairly intuitive. Consider its contrapositive: if a technology does not have an advantage is cost efficiency in any period, it will not be used; this is fairly obvious. Alternatively, if a technology is the most cost effective in every single period, it will be the only technology used. The second condition regarding output efficiency actually stems from the invertibility of $\xi$. If we did not have $\xi$ invertible, we either have at least one technology that does not produce in any period or we have at least one technology being a linear combination of the other technologies in terms of output. In the latter case, some technologies are not economical because their output can be replicated by the other technologies in a more cost effective way. That is, suppose $\xi$ consisted of 3 technologies but was of rank $2$. In this case, we may represent any of these technologies as a linear combination of the other two; call this combination the synthetic version of technology $i$. It is not possible for every synthetic technology $i$ to be more expensive than its original; for at least one $i$, we must have a synthetic technology being cheaper or equal in price to the actual technology. In the first case, synthetic technology $i$ is cheaper so we do not use the actual technology $i$. In the latter case, we may still eliminate technology $i$ or not use the other two technologies depending on the demand. Overall, in any case, $\xi$ not invertible means that at least one technology is not used. Finally, the demand condition is straightforward; even if a technology is optimal in a certain period, if consumers do not sufficiently demand electricity during that period, then there is little reason to use that technology.

We may also derive the comparative statics for this simplified scenario.  
\begin{proposition}
	Suppose we are not in an edge case, so that the conditions of Proposition 1 hold for each technology. The equilibrium quantity of a technology is increasing with its output efficiency and decreasing with its cost; at the same time, it is decreasing with the output efficiency of the other technologies and increasing with their cost. Also, suppose that some technology $i$ is the most cost effective in period $t$. Then, its equilibrium quantity is increasing with respect to the demand parameter $\alpha_t$ and decreasing with respect to the demand parameters of the other periods. Furthermore, again assuming technology $i$ is the most cost effective in period $t$, the comparative statics of $Z_t$ and $X_i$ are equivalent. 
\end{proposition}	
The comparative statics with respect to $X$ and its output efficiency and cost are not surprising. On the other hand, the statics for $Z$ may not be immediately obvious. Instead, they follow from the fact that we have $Z \equiv \xi_T X$. That is, suppose we have an arbitrary technology $i$ that is the most cost effective source of electricity in period $t$. If consumers demanded that 100\% of their energy arrive in period $t$, then relying solely on technology $i$ for energy would be the most economical solution. Consequently, it seems intuitive that the comparative statics of $X_i$ follow through to $Z_t$. This intuition just happens to apply even when other technologies are employed and there is demand in multiple periods. Similarly, the comparative statics for the share parameters of the utility function, $\alpha$, travel in the opposite direction. A rise in $\alpha_t$ would directly raise the optimal quantity of $Z_t$; hence, whichever technology is most cost effective at producing in period $t$ would be used more. We provide a more detailed and formal discussion of the comparative statics and edge cases in the appendix.

@@

\section{Results}
\label{sec:results}

@@

\subsection{Implications for Environmental Policy}

Overall, our results suggest that the elasticity of substitution between  renewable and fossil energy is not constant. Furthermore, it appears to fall with the level of intermittency. What implications does this result have?

\subsubsection{Revisiting Past Models}
